---
title: "'Ease of Doing Business' World Bank Ranking"
output: 
  html_document: default
---

```{r set_options, echo = FALSE, include = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)
```


```{r, include=FALSE}
setwd("C:/Users/pc/Documents/slowdata/R/doing_business")
```


Reading about the internazionalization of italian SMEs (Small and Medium Enterprises, < 250 employees) I discovered an interesting indicator provided by the World Bank. It is called 'Ease of doing business' and it is used to rank economies worldwide on their business-friendliness regulations (1=most business-friendly regulations).

Rankings are determined by sorting the aggregate 'distance to frontier' scores on 10 topics, each consisting of several indicators, giving equal weigth to each topic. More information on the methodologies can be found here http://www.doingbusiness.org/methodology.

In R there is a package to access World Bank data so you can avoid the hassle of downloading, importing, reformatting missing values and so on...


## World Bank Data

### Access data
```{r}
# access data
library(WDI)
WDIsearch('doing business') # when you don't know the exact name of an indicator you can search it by keyword
dat <- WDI(country = "all", 
           indicator = "IC.BUS.EASE.XQ", 
           start = 2016, 
           end = 2016)

dat[150,'country'] <- 'North Korea' # fixing no-utf characters in north korea name
```


Fine, but I'd like to have as well all the 10 components behind the index.

```{r}
WDIsearch('Dealing with') 
dim1 <- WDI(country = "all", 
           indicator = "IC.FRM.XQ", 
           start = 2016, 
           end = 2016) 
```

Apparently these dimesions are not available from API...
...but you can download them manually from [here](http://www.doingbusiness.org/data) in excel format.

```{r}
library(readxl)
ease_dim <- read_excel(path = "./data/Reports.xlsx", sheet = 1, col_names = TRUE, na = "", skip = 0)
keep <- grep(pattern = "Rank", x = names(ease_dim)) # keep only ranking variables
ease_dim <- ease_dim[,c(1,2,keep)]
```


### Clean data

Unfortunately there are no iso2c codes in here

```{r}
length(dat$country)
length(ease_dim$Economy) # probably regions (Europe, Middle-East, etc.) do not have ranking by single dimension

sum(ease_dim$Economy %in% dat$country) # 176 country names out of 212 are matched
ease_dim[!ease_dim$Economy %in% dat$country,'Economy'] 
```

Some countries in the two datasets do not match. This is because in the data downloaded from website there is also the index decomposed at the level of cities for some countries. We are not interested to that now. But there are also mismatches due to different names for the same country and we need to fix this.

```{r}
which(!ease_dim$Economy %in% dat$country)
ease_dim[which(!ease_dim$Economy %in% dat$country)[c(7, 8,9,10,15,18,21,29,30,36)], 'Economy'] <- c('Congo, Dem. Rep.', 'Congo, Rep.', "Cote d'Ivoire", 'Egypt, Arab Rep.', 'Iran, Islamic Rep.', 'North Korea', 'Micronesia, Fed. Sts.', 'Sao Tome and Principe', 'St. Kitts and Nevis', 'Yemen, Rep.')
```

Convert ranking columns to numeric

```{r}
for(i in 3:ncol(ease_dim)) {
  ease_dim[,i] <- as.numeric(unlist(ease_dim[,i]))
}
```


Now that names match we can merge

```{r}
dat2 <- merge(x = dat, y = ease_dim, by.x = 'country', by.y = 'Economy')
```


Are we getting the same data for the overall doingBusiness index??
```{r}
head(dat2[,c('IC.BUS.EASE.XQ', 'Rank')])
```

Well, actually not. Probably info from API are not updated, let's use what I found on the website (as of 29/03/2017).



## World admnistrative borders

Access borders from naturalearth website

```{r}
url <- "http://www.naturalearthdata.com/http//www.naturalearthdata.com/download/50m/cultural/ne_50m_admin_0_countries.zip"
tmp <- tempdir()
file <- basename(url)
download.file(url, file)
unzip(file, exdir = tmp)

library(rgdal)
library(sp)

countries <- readOGR(dsn = tmp, 
                     layer = "ne_50m_admin_0_countries", 
                     encoding = "UTF-8",
                     verbose = FALSE)

```


## Preparing plotting data

### Cleaning before matching world bank and natural Earth datasets

```{r}
sum(!dat2$iso2c %in% countries$iso_a2)
# dat2[which(!dat2$iso2c %in% countries$iso_a2), c('iso2c', 'country')] 
# countries@data[countries@data$name=='Kosovo','iso_a2']
countries@data$iso_new <- as.character(countries@data$iso_a2)
countries@data[countries@data$name=='Kosovo','iso_new'] <- 'XK' # Fixing Kosovo iso code...

# Fixing Somalia       
# countries@data[countries@data$name=='Somaliland','iso_a2']
# dat2[which(dat2$country=='Somalia'),'iso2c']
countries@data[countries@data$name=='Somaliland','iso_new'] <- 'SO' # Fixing Somalia
countries@data$iso_new2 <- as.factor(countries@data$iso_new)
```

### Merging

```{r}
plot.data <- merge(countries, 
                    dat2, 
                    by.x = "iso_new2", 
                    by.y = "iso2c",
                    sort = TRUE)
```



## Prepare layers

```{r}
library(leaflet)
library(RColorBrewer)
```

### Overall ranking
```{r}
q1 <- quantile(x = plot.data@data$Rank, probs = seq(0, 1, 0.1), na.rm = TRUE)
pal1 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q1)
popup1 <- paste0("<strong>Country: </strong>", plot.data$name, "<br><strong>", 'Ease of Doing Business Rank', ", ", 
                        as.character(2016), ": </strong>", plot.data[['Rank']])
```

### Starting a Business ranking
```{r}
q2 <- quantile(x = plot.data@data[,'Starting a business-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal2 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q2)
popup2 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Starting a Business Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Starting a business-Rank']])
```


### Dealing with Construction Permits-Rank ranking
```{r}
q3 <- quantile(x = plot.data@data[,'Dealing with Construction Permits-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal3 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q3)
popup3 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Dealing with Construction Permits-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Dealing with Construction Permits-Rank']])
```

### Dealing with Construction Permits-Rank
```{r}
q4 <- quantile(x = plot.data@data[,'Getting Electricity-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal4 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q4)
popup4 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Getting Electricity-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Getting Electricity-Rank']])
```


### Registering Property-Rank
```{r}
q5 <- quantile(x = plot.data@data[,'Registering Property-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal5 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q5)
popup5 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Getting Electricity-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Registering Property-Rank']])
```


### Getting Credit-Rank
```{r}
q6 <- quantile(x = plot.data@data[,'Getting Credit-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal6 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q6)
popup6 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Getting Credit-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Getting Credit-Rank']])
```


### Protecting Minority Investors-Rank
```{r}
q7 <- quantile(x = plot.data@data[,'Protecting Minority Investors-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal7 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q7)
popup7 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Protecting Minority Investors-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Protecting Minority Investors-Rank']])
```



### Paying Taxes-Rank
```{r}
q8 <- quantile(x = plot.data@data[,'Paying Taxes-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal8 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q8)
popup8 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Paying Taxes-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Paying Taxes-Rank']])
```



### Trading Across Borders-Rank
```{r}
q9 <- quantile(x = plot.data@data[,'Trading Across Borders-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal9 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q9)
popup9 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Trading Across Borders-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Trading Across Borders-Rank']])

```



### Enforcing Contracts-Rank
```{r}
q10 <- quantile(x = plot.data@data[,'Enforcing Contracts-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
pal10 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = q10)
popup10 <- paste0("<strong>Country: </strong>", 
                 plot.data$name, 
                 "<br><strong>", 
                 'Enforcing Contracts-Rank', 
                 ", ", 
                 as.character(2016), 
                 ": </strong>", 
                 plot.data[['Enforcing Contracts-Rank']])
```



### Resolving Insolvency-Rank
```{r}
q11 <- quantile(x = plot.data@data[,'Resolving Insolvency-Rank'], probs = seq(0, 1, 0.1), na.rm = TRUE)
q11 # breaks are not unique
manual_q11 <- c(1,21.2,40.2, 58.8, 77.4, 96.0, 115.6, 134.2, 152.8, 160, 169.0)
pal11 <- colorBin(rev(brewer.pal(n = 9, name = "Greens")), bins = manual_q11)
popup11 <- paste0("<strong>Country: </strong>", 
                  plot.data$name, 
                  "<br><strong>", 
                  'Resolving Insolvency-Rank', 
                  ", ", 
                  as.character(2016), 
                  ": </strong>", 
                  plot.data[['Resolving Insolvency-Rank']])

```


### Prepare map background
```{r}
stamen_tiles <- "http://{s}.tile.stamen.com/toner-lite/{z}/{x}/{y}.png"
stamen_attribution <- 'Map tiles by <a href="http://stamen.com">Stamen Design</a>, under <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a>. Data by <a href="http://openstreetmap.org">OpenStreetMap</a>, under <a href="http://www.openstreetmap.org/copyright">ODbL</a>.'
```


## Plot

```{r}
DoingBusinessMap <-
leaflet(data = plot.data) %>%
  addTiles(urlTemplate = stamen_tiles,  
           attribution = stamen_attribution) %>%
  setView(0, 0, zoom = 3) %>%
  addPolygons(fillColor = ~pal1(plot.data[['Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup1,
              group = "<span style='color: #7f0000; font-size: 11pt'><strong>Ease of Doing Business Rank</strong></span>") %>%
  addPolygons(fillColor = ~pal2(plot.data[['Starting a business-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup2,
              group = "Starting a Business-Rank") %>%
  addPolygons(fillColor = ~pal3(plot.data[['Dealing with Construction Permits-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup3,
              group = "Dealing with Construction Permits-Rank") %>%
  addPolygons(fillColor = ~pal4(plot.data[['Getting Electricity-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup4,
              group = "Getting Electricity-Rank") %>%
  addPolygons(fillColor = ~pal5(plot.data[['Registering Property-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup5,
              group = "Registering Property-Rank") %>%
  addPolygons(fillColor = ~pal6(plot.data[['Getting Credit-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup6,
              group = "Getting Credit-Rank") %>%
  addPolygons(fillColor = ~pal7(plot.data[['Protecting Minority Investors-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup7,
              group = "Protecting Minority Investors-Rank") %>%
  addPolygons(fillColor = ~pal8(plot.data[['Paying Taxes-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup8,
              group = "Paying Taxes-Rank") %>%
  addPolygons(fillColor = ~pal9(plot.data[['Trading Across Borders-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup9,
              group = "Trading Across Borders-Rank") %>%
  addPolygons(fillColor = ~pal10(plot.data[['Enforcing Contracts-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup10,
              group = "Enforcing Contracts-Rank") %>%
  addPolygons(fillColor = ~pal11(plot.data[['Resolving Insolvency-Rank']]), 
              fillOpacity = 0.8, 
              color = "#BDBDC3", 
              weight = 1, 
              popup = popup11,
              group = "Resolving Insolvency-Rank") %>%
  addLayersControl(
    baseGroups = c("<span style='color: #7f0000; font-size: 11pt'><strong>Ease of Doing Business Rank</strong></span>", 
                   "Starting a Business-Rank",
                   "Dealing with Construction Permits-Rank",
                   "Getting Electricity-Rank",
                   "Registering Property-Rank",
                   "Getting Credit-Rank",
                   "Protecting Minority Investors-Rank",
                   "Paying Taxes-Rank",
                   "Trading Across Borders-Rank",
                   "Enforcing Contracts-Rank",
                   "Resolving Insolvency-Rank"
    ),
    options = layersControlOptions(collapsed = FALSE)) %>%
  addLegend(position = 'bottomleft', 
            colors = rev(brewer.pal(n = 9, name = 'Greens')), 
            labels = paste0(rep("<span style='font-size: 8pt'>", 9), seq(10,90,10), rep("<sup>th</sup></span>",9)),  
            opacity = 0.8,      ##transparency again
            title = "<span style='font-size: 9pt'> Top-Percentiles</span>")
```



## Saving plot
```{r}
library(htmlwidgets)
# saveWidget(DoingBusinessMap, "doingbiz.html", selfcontained = FALSE)
```